---
title: "Wordle on keyboard"
subtitle: "Coding out loud, Ep 4"
author: "Mine Çetinkaya-Rundel + Martha Aboagye"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(ggkeyboard) # https://github.com/sharlagelfand/ggkeyboard
```

## Inspiration

[English keyboard heatmap](https://twitter.com/monachalabi/status/772102591694209024) by [Mona Chalabi](https://twitter.com/MonaChalabi)

## Task 1: Find the first letter of Wordle words

From https://github.com/tabatkins/wordle-list.

```{r}
wordle <- read_csv("https://raw.githubusercontent.com/tabatkins/wordle-list/main/words", col_names = "word")
```

```{r}
wordle <- wordle %>%
  mutate(first_letter = str_sub(word, start = 1, end = 1))
```

```{r}
wordle_counts <- wordle %>%
  count(first_letter, sort = TRUE)
```


## Task 2: Draw a keyboard

```{r}
ggkeyboard(tkl, layout = "iso")
```

## Task 3: Color keyboard keys based on frequency of first letter

### Color options

dark blue - "#2D669B"
light blue - "#359CA8"
green - "#B8C222"
yellow - "#EEDD49"
orange - "#F2BC4A"
dark orange - "#FB704A"
red - "#BE372C"

### Color scale

Let's do some exploratory analysis to determine a scale based on the distribution of `n`.

```{r}
wordle_counts %>%
  summarise(
    min = min(n),
    mean = mean(n),
    sd = sd(n),
    max = max(n),
    q20 = quantile(n, 0.20),
    q40 = quantile(n, 0.40),
    q60 = quantile(n, 0.60),
    q80 = quantile(n, 0.80)
  )

ggplot(wordle_counts, aes(x = n)) +
  geom_histogram(binwidth = 100)

ggplot(wordle_counts, aes(x = first_letter, y = n)) +
  geom_col() +
  geom_hline(yintercept = 80)
```

Based on these, we can use the following cutoffs.

- dark blue   - "#2D669B" - 0
- light blue  - "#359CA8" - 20
- green       - "#B8C222" - 189
- yellow      - "#EEDD49" - 325
- orange      - "#F2BC4A" - 595
- dark orange - "#FB704A" - 736
- red         - "#BE372C" - > 1000

Then, assign the colors to the letters.

```{r}
wordle_counts <- wordle_counts %>%
  mutate(
    color = case_when(
      n == 0                  ~ "#2D669B",
      n <=  20                ~ "#359CA8",
      n > 20  & n <= 189      ~ "#B8C222",
      n > 189 & n <= 325      ~ "#EEDD49",
      n > 325 & n <= 595      ~ "#F2BC4A",
      n > 595 & n <= 1000     ~ "#FB704A",
      n > 1000  ~ "#BE372C"
      ),
    first_letter = str_to_upper(first_letter)
  ) %>%
  arrange(first_letter)
```

Experiment a bit with the `ggkeyboard()` function to see how colors get assigned to letters.

```{r}
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) %>%
  highlight_keys(
    keys = c("Z", "X", "C", "V"), 
    fill = c("red", "pink", "orange", "blue"), 
    alpha = 0.7, 
    colour = NA
    )

ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) %>%
  highlight_keys(
    keys = c("C", "N", "S"), 
    fill = c("red", "pink", "orange"), 
    alpha = 0.7, 
    colour = NA
    )
```

Looks like we need to order the letters in the order they appear on the keyboard, so let's do that.

```{r}
keys <- read_csv("https://raw.githubusercontent.com/sharlagelfand/ggkeyboard/061f0d9084d76d576200e81439311983e1d50d93/data-raw/sixty_percent.csv")

wordle_counts_ordered <- keys %>% 
  filter(key_label %in% LETTERS) %>%
  select(key) %>%
  left_join(wordle_counts, by = c("key" = "first_letter"))
```

And finally we can assign the colors to the keys.

```{r}
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) %>%
  highlight_keys(
    keys = wordle_counts_ordered$key, 
    fill = wordle_counts_ordered$color, 
    alpha = 0.7, 
    colour = NA
    )
```

Remaining to do items: change the background of the keyboard and non-letter keys and add a legend.
